Recalibrating Gravitational Wave Phenomenological Waveform Model

### 2023-06-27 src/tex/ms.tex
fixing minor typos

Many data analysis tasks in gravitational wave (GW) astrophysics, such
as match filtering~\citep{owen1996search, owen1999matched} and parameter 
estimation~\citep{Dax:2021tsq, Christensen:2022bxb, dynesty, Islam:2022afg, 
Romero-Shaw:2020owr, zackay2018relative, Ashton:2018jfp}, rely upon accurate waveform
models. Since the generation of numerical relativity (NR) waveforms is
prohibitively expensive, the community has constructed waveform approximants
that can be evaluated much faster. There are three families of commonly used GW
approximants: the effective-one-body (EOB) \citep{ossokine2020multipolar,
cotesta2020frequency, taracchini2014effective}, NR
surrogate \citep{islam2022surrogate, varma2019surrogate, varma2019surrogate2},
and phenomenological (Phenom) models \citep{husa2016frequency,
khan2016frequency, garcia2020multimode, pratten2021computationally}. While the
detailed construction of each model is different, they all have a set of
internal parameters that can be calibrated to NR waveforms. The quality of the
waveform model is therefore determined by the ansatz used and the accuracy of the
calibrated parameters.

@@

% Put more focus on the joint optimization instead of having a set of more accurate 
% coefficients, since the method is the main point
In this paper, we investigate the possibility of further improving the accuracy
of a waveform model, IMRPhenomD \citep{khan2016frequency, husa2016frequency}, by jointly optimizing all the
calibration coefficients for given a set of NR waveforms. Using a similar set of NR
waveforms to those used in \citep{khan2016frequency, husa2016frequency},
we demonstrate that one can improve the match between IMRPhenomD and NR
waveforms over a decently sized parameter space, up to a mass ratio $q=8$. We additionally
explore how different parts of the source parameter space (e.g. the primary and
secondary spins) respond to the optimization procedure by optimizing the waveform
separately for different regions. This can help in understanding
whether the waveform model ansatz performs equally well in different regions of the
parameter space.

@@

In this section, we first briefly review the construction of the IMRPhenomD model and discuss
how the calibration parameters enter the waveform.
We then outline the mismatch and how it can be used as a loss function. 
Finally we discuss the gradient descent algorithm and our stopping criterion.

@@

Aligned-spin, frequency-domain waveform models (such as IMRPhenomD) can be written as a
combination of amplitude and phase functions ($A$ and $\phi$ respectively):
\begin{align}\label{eq:}
	h(f,\theta,\Lambda) = A(f,\theta,\Lambda)e^{-i\phi(f,\theta,\Lambda)}\,,
\end{align}
where $f$ is the frequency, $\theta$ are the intrinsic parameters of the binary, and $\Lambda$ is a set of additional
parameters which will be discussed below. 
The phase and amplitude functions are then split into three sections which represent the
inspiral, intermediate, and merger-ringdown (MR) parts of the waveform. 
During inspiral, $A$ and $\phi$ are known analytically from post-Newtonian (PN) theory;
IMRPhenomD uses the TaylorF2 model \citep{Buonanno:2009zt, Arun:2004hn} which is known up to 3.5PN order.
To model the intermediate and MR regions, IMRPhenomD (and all waveforms in the IMRPhenom family)
uses a series of parameterizations\footnote{
	The parameterizations for both the amplitude and phase functions can be found in \citep{khan2016frequency}.
} 
which depend purely on $\Lambda$ and can be calibrated to numerical relativity (NR) simulations.
The three sections are then \textit{stitched} together using step functions.
Importantly, the parameterizations are chosen such that they can be made $\mathcal{C}^1$ continuous at the
boundary between each section.

In practice the $\Lambda$ parameters are fit for each section independently, i.e., intermediate coefficients are fit whilst ignoring the MR region.
Finally, to map the grid of tuned $\Lambda$ parameters back to the intrinsic parameter space, IMRPhenomD uses the polynomial function:
\begin{align} \label{eq:Lambda}
	\Lambda^i&=\lambda_{00}^i+\lambda_{10}^i\eta \nonumber \\
	&+(\chi_{\mathrm{PN}}-1)(\lambda_{01}^i+\lambda_{11}^i\eta+\lambda_{21}^i\eta^2) \nonumber \\ 
	&+(\chi_{\mathrm{PN}}-1)^2(\lambda_{02}^i+\lambda_{12}^i\eta+\lambda_{22}^i\eta^2) \nonumber \\
	&+(\chi_{\mathrm{PN}}-1)^3(\lambda_{03}^i+\lambda_{13}^i\eta+\lambda_{23}^i\eta^2)\,,
\end{align}
where the $\lambda$'s are the fitting coefficients we are going to optimize below, $\eta$ is
the symmetric mass ratio, and $\chi_{\mathrm{PN}}$ is the post-Newtonian spin
parameter, which is defined as 
\begin{align}
	\chi_{\mathrm{PN}}=\frac{m_1\chi_1+m_2\chi_2}{m_1+m_2}-\frac{38\eta}{113}(\chi_1+\chi_2)\,.
\end{align}
Here, $m_{1,2}$ and $\chi_{1,2}$ are the primary and secondary mass and spin,
respectively.

@@

In order to optimize the coefficients, we need to define a loss function that
quantifies the difference between the waveform model and the target NR simulations which we want to match.
Here we adopt a quantity commonly used in GW physics called the \textit{mismatch} function \citep{owen1996search, husa2016frequency}. 
It is defined as
\begin{align} \label{eq:mismatch}
	\mathcal{M}(h_1, h_2)=1-\max_{t_0, \phi_0}\langle \hat{h}_1, \hat{h}_2\rangle,
\end{align}
where $h_{1,2}$ are the two GW waveforms we are comparing, and $t_0$ and $\phi_0$
are a relative time and phase shift respectively. 
The inner product, $\langle h_1, h_2 \rangle$, is defined as 
\begin{align}\label{eq:inner_prod}
	\langle h_1, h_2 \rangle = 4\Re\int_{f_{\mathrm{min}}}^{f_{\mathrm{max}}}\frac{h_1(f)h_2^{\ast}(f)}{S_n(f)}\,df,
\end{align}
where $\hat{h}=h/\sqrt{\langle h, h \rangle}$ is the normalized GW strain,
$S_n(f)$ is the power spectral density (PSD), and $f_{\mathrm{max}}$ ($f_{\mathrm{min}}$) is
the maximum (minimum) frequency for the integration.
We note here that the mismatch can be viewed as the mean square error (MSE) between the
two waveforms.

Since we wish to optimize the model over the whole parameter space,
we need to compare multiple model generated waveforms with NR waveforms.
However, the mismatch is only defined for two input waveforms at a particular set of intrinsic parameters.
We therefore define the loss function as an average of training waveforms in two ways,
the simple average of mismatches and the normalized average of mismatches,  
\begin{align}\label{eqn:loss}
	\mathcal{L}_{\mathrm{ave}}&=\frac{1}{N}\sum_{i=1}^N\mathcal{M}_i \\
	\mathcal{L}_{\mathrm{norm}}&=\frac{1}{N}\sum_{i=1}^N\frac{\mathcal{M}_i}{\mathcal{M}_{i,\mathrm{ini}}},
\end{align}	
where $\mathcal{M}_i$ represents the mismatch of an individual training waveform,
$\mathcal{M}_{i,\mathrm{ini}}$ represents the initial mismatch of the individual
training waveform, and $N$ is the total number of training waveforms.

The two loss functions are chosen to give different 
behavior during the optimization process.
The simple average, $\mathcal{L}_{\mathrm{mean}}$, serves as the simplest choice of loss function,
but is prone to be dominated by a single point in parameter space with a large mismatch.
Other points with smaller mismatches would be insignificant comparatively, and might not be able 
to improve under such a loss function.
Alternatively, the normalized average, $\mathcal{L}_{\mathrm{norm}}$, eliminates the aforementioned issue
by encouraging the waveform to improve at each training point at a similar rate.
The ratio in $\mathcal{L}_{\mathrm{norm}}$ will therefore remain approximately the same for each training point.
Conversely, $\mathcal{L}_{\mathrm{mean}}$ allows the loss function to automatically adjust
and preferentially optimize the largest mismatches, encouraging the waveform to have similar mismatches everywhere.
In this paper, we show the results of using both loss functions 
and examine the differences between them.

@@

To compute the loss functions, we have to choose NR waveforms for calculating the mismatch.
Originally, 19 NR simulations were used to calibrate IMRPhenomD \citep{khan2016frequency}; 9 of these 
are from the SXS catalog \citep{boyle2019sxs} and 10 BAM simulations. 
As BAM waveforms are not publicly available, we cannot use a training set identical to the original work.
Instead, we take the same nine waveforms from the SXS catalog plus two additional waveforms which closely mimic
two of the low mass ratio BAM simulations. 
The remaining BAM simulations have no close SXS counterparts, and are therefore not included in the training set.
Our main results therefore uses 11 NR waveforms for calibration which are listed in Tab.~\ref{tab:q148}.
Additional NR waveforms which are used for further calibration are listed in Tab.~\ref{tab:q1248}.

The training set has a maximum mass ratio of eight due to the lack 
of high mass ratio simulations in the SXS catalog.
In fact, the SXS catalog only has NR waveforms with $q\leq10$.
Nevertheless, we are interested in the behavior of IMRPhenomD model with small $q$,
as most BBH events observed by LIGO and Virgo have $q\leq8$.

The SXS NR waveforms are given as time-series strain.
Since IMRPhenomD is modeled in the frequency domain, we need to Fourier transform the simulation results
in order to compute the mismatch, \eqref{eq:mismatch}.
For this, we taper the time-series using Tukey window~\citep{usman2016pycbc}\footnote{
	Specifically, we choose $\alpha=2t_{\mathrm{RD}}/T$, where
	$t_{\mathrm{RD}}$ is the duration of ringdown (maximum amplitude to the end of the strain) and $T$ is the duration of the
	entire GW strain. \te{What is $t_{\mathrm{RD}}$? and how is it computed?}
} before using standard FFT routines to compute the Fourier transform.

In addition to choosing the NR waveforms for the training set,
one needs to choose a relevant PSD for mismatch.
We have opted to use a flat PSD for the mismatch calculation,
as it provides results that are independent of the detector sensitivity and mass scale.
The use of a flat PSD ensures that the improvement in accuracy is due mainly
to the difference in high-dimensional fitting.
Additionally, we are interested in examining the effect of introducing a detector PSD
on the optimization process. For this, we have chosen the
zero-detuned high-power (\zdethp) noise PSD \citep{aasi2015advanced}.
Since the total mass of the system scales with the frequency of the waveform,
we must choose a corresponding mass scale to match the frequency range of our noise PSD.
To demonstrate the effect of introducing a detector specific PSD,
we selected an arbitrary mass scale of $M=50M_{\odot}$, as binaries
of this mass are commonly observed by the LIGO and Virgo detectors~\citep{gwtc1, gwtc2, gwtc21, gwtc3}.

@@

With the loss function evaluated, we apply gradient descent to optimize the tunable
coefficients as shown in Algorithm~\ref{alg:gradient}. We take $\lambda_i$ to be the 
original coefficients given in \citep{khan2016frequency} because they likely lie in the
neighborhood of the minimum that we wish to find. 
We fix our learning rate, $\alpha$, to be $=10^{-6}$, which is small enough to ensure we 
don't move far from the minimum.
Finally, we stop the optimization when the validation loss stops decreasing 
(see Sec.~\ref{sec:result} for description of the validation set).
This can be seen in Fig.~\ref{fig:loss} at around $12000$ iterations.

\begin{figure}
	\script{loss.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/loss.pdf}
	\caption{Loss functions against the number of iterations. The vertical
	dotted line indicates minimum of the validation loss, which is where we stop
	the optimization.}
	\label{fig:loss}
\end{figure}

@@

\begin{figure}[t]
	\script{intrin_space.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/intrin_space.pdf}
	\caption{Distribution of training and validation NR waveforms in the
	$q-\chi_{PN}$ space. The blue stars represent the waveforms used to compute
	the training loss and the orange dots represent the waveforms used for
	validation.}
	\label{fig:intrin_space}
\end{figure}

\begin{table}[t]
	\centering
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
		\midrule\midrule SXS:BBH:0156 & 1.0 & -0.95    & -0.95    \\
		SXS:BBH:0151 & 1.0 & -0.60    & -0.60    \\
		SXS:BBH:0001 & 1.0 &  0.00    &  0.00    \\
		SXS:BBH:0152 & 1.0 &  0.60    &  0.60    \\
		SXS:BBH:0172 & 1.0 &  0.98    &  0.98    \\
		SXS:BBH:1418 & 4.0 & -0.40    & -0.50    \\
		SXS:BBH:0167 & 4.0 &  0.00    &  0.00    \\
		SXS:BBH:1417 & 4.0 &  0.40    &  0.50    \\
		SXS:BBH:0064 & 8.0 & -0.50    & -0.46    \\
		SXS:BBH:0063 & 8.0 &  0.00    &  0.00    \\
		SXS:BBH:0065 & 8.0 &  0.50    &  0.46    \\ \midrule\bottomrule
	\end{tabularx}
	\caption{List of waveforms used to recalibrate the model. The mass ratio is defined as
	$q=m_1/m_2\geq 1$ and the spins are denoted by $\chi_{1,2}$. Out of the 11 waveforms listed
	here, two SXS waveforms are analogues to two of the BAM NR simulations used in the original calibration (SXS:BBH:1417 as A7 and SXS:BBH:1418 as A9 in \citep{khan2016frequency}). The rest of the SXS waveforms are also used in the original IMRPhenomD calibration.}
	\label{tab:q148}
\end{table}
\begin{table}[t]
	\centering
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
		\midrule\midrule SXS:BBH:0234 & 2.0 & -0.85    & -0.85    \\
		SXS:BBH:0235 & 2.0 & -0.60    & -0.60    \\
		SXS:BBH:0169 & 2.0 & 0.00     & 0.00     \\
		SXS:BBH:0256 & 2.0 & 0.60     & 0.60     \\
		SXS:BBH:0257 & 2.0 & 0.85     & 0.85     \\ \midrule\bottomrule
	\end{tabularx}
	\caption{Additional NR waveforms used in further recalibration.}
	\label{tab:q1248}
\end{table}

@@

To evaluate how well the optimization procedure generalizes to waveforms that are
not in the training set, we evaluate the mismatch between the fine-tuned model
and an additional 536 NR waveforms in the SXS catalog i.e., the validation set.
We select waveforms that share the same part of the parameter space with the training
set, i.e., waveforms with negligible eccentricity (${e<2\times10^{-3}}$)
and precession (${\chi_{x,y}<5\times10^{-3}}$). Figure~\ref{fig:intrin_space}
shows how the training and validation waveforms are distributed in the
$q-\chi_{\rm PN}$ space.

To illustrate the effect of optimization on an individual waveform level, in Fig.
\ref{fig:0154} we plot the phase and amplitude of a particular waveform before and after optimization together with
the NR waveform taken directly from the SXS catalog. In the bottom panel one can see that,
compared to the original IMRPhenomD waveform, the optimized waveform has smaller residuals both in amplitude and
phase, particularly in the inspiral region where the amplitude displays a
$50\%$ reduction in error. For a fair comparison, we selected the SXS:BBH:0154 NR waveform
which was also used in \citep{khan2016frequency} to validate the original waveform.

With the purpose of improving downstream tasks such as parameter estimation in
mind, the more relevant metric of improvement is the distribution of improvement
in mismatch over the entire validation dataset. Figure~\ref{fig:q148} shows the
distribution of log-mismatches for the validation waveforms before and after the
optimization procedure. Here we show results using a constant PSD in our loss
function. One can see the distribution of the optimized waveform is skewed toward lower mismatches,
with the peak of the distribution being shifted by approximately an order of
magnitude, and the median mismatch being reduced by 50\% (see vertical dotted lines). When using
$\mathcal{L}_{\mathrm{norm}}$, we observe a less pronounced improvement with a 22.9\%
decrease in the median.

@@

Note that the performance of the IMRPhenomD model was initially tested using the
{\zdethp} weighted mismatch.\footnote{
	Note, however, that the mismatch was never directly used during the calibration process~\citep{khan2016frequency, khan2019phenomenological}.
} We would therefore like to examine whether using the {\zdethp} PSD in our loss function
could lead to an improved to mismatch. Performing the optimization again, we find no significant difference
between the results using the two PSDs in the distribution of mismatches.

@@

To investigate the performance of recalibration over the source parameter space,
we plot the improvement of the log-mismatch as a function of the parameter space
$q-\chi_{\mathrm{PN}}$ in Fig. \ref{fig:ps_q148_qchi}. Red points indicate that the waveform
is improved by the optimization procedure, while blue points indicate that the waveform mismatch
\textit{increases} during optimization.
We can see that waveforms with $q\leq4$ show the most consistent average improvement.
This is likely due to the better coverage of training waveforms in that part of the parameter space (see Fig.~\ref{fig:intrin_space}).
On the spin axis, we can see that the waveforms with
$\chi_{\textrm{PN}}\sim0$ show the most consistent improvement. 
When we move away from $\chi_{\textrm{PN}}\sim0$, the improvement fluctuates but exhibits an overall trend. 
This is particularly true in the $q\leq4$ region, where we see a consistent
improvement of the waveform for $\chi_{\textrm{PN}}<0$. 
We also plot the parameter space $\chi_1 - \chi_2$ in Fig.~\ref{fig:ps_q148}. 
Points along the diagonal axis, $\chi_1\sim\chi_2$, show good mismatch improvements as discussed
above. Meanwhile, the top-left and bottom-right regions respond to the optimization differently. In the top-left
region, the waveform generally improves after optimization. However,
in the bottom-right region, the waveform does not improve after optimization.

>>>>

Many data analysis tasks in gravitational wave (GW) astrophysics, such
as match filtering~\citep{owen1996search, owen1999matched} and parameter 
estimation~\citep{Dax:2021tsq, Islam:2022afg, zackay2018relative, Ashton:2018jfp}, rely upon accurate waveform
models. Since the generation of numerical relativity (NR) waveforms is
prohibitively expensive, the community has constructed waveform approximants
that can be evaluated much faster. There are three families of commonly used GW
approximants: the effective-one-body (EOB) \citep{ossokine2020multipolar,
cotesta2020frequency, taracchini2014effective}, NR
surrogate \citep{islam2022surrogate, varma2019surrogate, varma2019surrogate2},
and phenomenological (Phenom) models \citep{husa2016frequency,
khan2016frequency, garcia2020multimode, pratten2021computationally}. While the
detailed construction of each model is different, they all have a set of
internal parameters that can be calibrated to NR waveforms. The quality of the
waveform model is therefore determined by the ansatz used and the accuracy of the
calibrated parameters.

@@

% Put more focus on the joint optimization instead of having a set of more accurate 
% coefficients, since the method is the main point
In this paper, we investigate the possibility of further improving the accuracy
of a waveform model, IMRPhenomD \citep{khan2016frequency, husa2016frequency}, by jointly optimizing all the
calibration coefficients for given a set of NR waveforms. Using a similar set of NR
waveforms to those used in \citep{khan2016frequency, husa2016frequency},
we demonstrate that one can improve the match between IMRPhenomD and NR
waveforms over a decently sized parameter space, up to a mass ratio $q=8$. We additionally
explore how different parts of the source parameter space (e.g. the primary and
second spins) respond to the optimization procedure by optimizing the waveform
separately for different regions. This can help in understanding
whether the waveform model ansatz performs equally well in different regions of the
parameter space.

@@

In this section, we first briefly review the construction of the IMRPhenomD model and discuss
how the calibration parameters enter the waveform.
We then outline the mismatch and how it can be used as a loss fucntion. 
Finally we discuss the gradient descent algorithm and our stopping criterion.

@@

Aligned-spin, frequency-domain waveform models (such as IMRPhenomD) can be written as a
combination of amplitude and phase functions ($A$ and $\phi$ respectively):
\begin{align}\label{eq:}
	h(f,\theta,\Lambda) = A(f,\theta,\Lambda)e^{-i\phi(f,\theta,\Lambda)}\,,
\end{align}
where $f$ is the frequnecy, $\theta$ are the intrinsic parameters of the binary, and $\Lambda$ is a set of additional
parameters which will be discussed below. 
The phase and amplitude functions are then split into three sections which represent the
inspiral, intermediate, and merger-ringdown (MR) parts of the waveform. 
During inspiral, $A$ and $\phi$ are known analytically from post-Newtonian (PN) theory;
IMRPhenomD uses the TaylorF2 model \citep{Buonanno:2009zt, Arun:2004hn} which is known up to 3.5PN order.
To model the intermediate and MR regions, IMRPhenomD (and all waveforms in the IMRPhenom family)
uses a series of parameterizations\footnote{
	The parameterizations for both the ampltude and phase functions can be found in \citep{khan2016frequency}.
} 
which depend purely on $\Lambda$ and can be calibrated to numerical relativity (NR) simulations.
The three sections are then \textit{stictched} together using step functions.
Importantly, the parameterizations are chosen such that they can be made $\mathcal{C}^1$ continuous at the
boundary between each section.

In practice the $\Lambda$ parameters are fit for each section independently i.e., intermediate coefficients are fit whilst ignoring the MR region.
Finally, to map the grid of tuned $\Lambda$ parameters back to the intrinsic parameter space, IMRPhenomD uses the polynomial function:
\begin{align} \label{eq:Lambda}
	\Lambda^i&=\lambda_{00}^i+\lambda_{10}^i\eta \nonumber \\
	&+(\chi_{\mathrm{PN}}-1)(\lambda_{01}^i+\lambda_{11}^i\eta+\lambda_{21}^i\eta^2) \nonumber \\ 
	&+(\chi_{\mathrm{PN}}-1)^2(\lambda_{02}^i+\lambda_{12}^i\eta+\lambda_{22}^i\eta^2) \nonumber \\
	&+(\chi_{\mathrm{PN}}-1)^3(\lambda_{03}^i+\lambda_{13}^i\eta+\lambda_{23}^i\eta^2)\,,
\end{align}
where the $\lambda$'s are the fitting coefficients we are going to optimize below, $\eta$ is
the symmetric mass ratio, and $\chi_{\mathrm{PN}}$ is the post-Newtonian spin
parameter, which is defined as 
\begin{align}
	\chi_{\mathrm{PN}}=\frac{m_1\chi_1+m_2\chi_2}{m_1+m_2}-\frac{38\eta}{113}(\chi_1+\chi_2)\,.
\end{align}
Here, $m_{1,2}$ and $\chi_{1,2}$ are the primary and secondary mass and spin,
respectively.

@@

In order to optimize the coefficients, we need to define a loss function that
quantifies the difference between the waveform model and the target NR simulations which we want to match.
Here we adopt a quantity commonly used in GW physics called the \textit{mismatch} function \citep{Owen:1995tm,husa2016frequency}. 
It is defined as
\begin{align} \label{eq:mismatch}
	\mathcal{M}(h_1, h_2)=1-\max_{t_0, \phi_0}\langle \hat{h}_1, \hat{h}_2\rangle,
\end{align}
where $h_{1,2}$ are the two GW waveforms we are comparing, and $t_0$ and $\phi_0$
are a relative time and phase shift respectively. 
The inner product, $\langle h_1, h_2 \rangle$, is defined as 
\begin{align}\label{eq:inner_prod}
	\langle h_1, h_2 \rangle = 4\Re\int_{f_{\mathrm{min}}}^{f_{\mathrm{max}}}\frac{h_1(f)h_2^{\ast}(f)}{S_n(f)}\,df,
\end{align}
where $\hat{h}=h/\sqrt{\langle h, h \rangle}$ is the normalized GW strain,
$S_n(f)$ is the power spectral density (PSD), and $f_{\mathrm{max}}$ ($f_{\mathrm{min}}$) is
the maximum (minimum) frequency for the integration.
We note here that the mismatch can be viewed as the mean square error (MSE) between the
two waveforms.

Since we wish to optimize the model over the whole parameter space,
we need to compare multiple model generated waveforms with NR waveforms.
However, the mismatch is only defined for two input waveforms at a particular set of intrinsic parameters.
We therefore define the loss function as an average of training waveforms in two ways,
the simple average of mismatches and the normalized average of mismatches,  
\begin{align}\label{eqn:loss}
	\mathcal{L}_{\mathrm{ave}}&=\frac{1}{N}\sum_{i=1}^N\mathcal{M}_i \\
	\mathcal{L}_{\mathrm{nl}}&=\frac{1}{N}\sum_{i=1}^N\frac{\mathcal{M}_i}{\mathcal{M}_{i,\mathrm{ini}}},
\end{align}	
where $\mathcal{M}_i$ represents the mismatch of an individual training waveform,
$\mathcal{M}_{i,\mathrm{ini}}$ represents the initial mismatch of the individual
training waveform, and $N$ is the total number of training waveforms.

The two loss functions are chosen to give different 
behaviour during the optimization process.
The simple average, $\mathcal{L}_{\mathrm{mean}}$, serves as the simplest choice of loss function,
but is prone to be dominated by a single point in parameter space with a large mismatch.
Other points with smaller mismatches would be insignificant comparatively, and might not be able 
to improve under such a loss function.
Alternatively, the normalized average, $\mathcal{L}_{\mathrm{norm}}$, eliminates the aforementioned issue
by encouraging the waveform to improve at each training point at a similar rate.
The ratio in $\mathcal{L}_{\mathrm{norm}}$ will therefore remain approximately the same for each training point.
Conversely, $\mathcal{L}_{\mathrm{mean}}$ allows the loss function to automatically adjust
and preferencially optimize the largest mismatches, encouraging the waveform to have similar mismatches everywhere.
In this paper, we show the results of using both loss functions 
and examine the differences between them.

@@

To compute the loss functions, we have to choose NR waveforms for calculating the mismatch.
Originally, 19 NR simulations were used to calibrate IMRPhenomD \citep{khan2016frequency}; nine of these 
are from the SXS catalog \citep{boyle2019sxs} and 10 BAM simulations. 
As BAM waveforms are not publicly available, we cannot use a training set identical to the original work.
Instead, we take the same nine waveforms from the SXS catalog plus two additional waveforms which closely mimic
two of the low mass ratio BAM simulations. 
The remaining BAM simulations have no close SXS counterparts, and are therefore not included in the training set.
Our main results therefore uses 11 NR waveforms for calibration which are listed in Tab.~\ref{tab:q148}.
Additional NR waveforms which are used for further calibration are listed in Tab.~\ref{tab:q1248}.

The training set has a maximum mass ratio of eight due to the lack 
of high mass ratio simulations in the SXS catalog.
In fact, the SXS catalog only has NR waveforms with $q\leq10$.
Nevertheless, we are interested in the behavior of the IMRPhenomD model with small $q$,
as most BBH events observed by LIGO and Virgo have $q\leq8$.

The SXS NR waveforms are given as time-series strain.
Since IMRPhenomD is modeled in the frequency domain, we need to Fourier transform the simulation results
in order to compute the mismatch, \eqref{eq:mismatch}.
For this, we taper the time-series using Tukey window~\citep{tukey}\footnote{
	Specifically, we choose $\alpha=2t_{\mathrm{RD}}/T$, where
	$t_{\mathrm{RD}}$ is the duration of ringdown and $T$ is the duration of the
	entire GW strain. \te{What is $t_{\mathrm{RD}}$? and how is it computed?}
} before using standard routines to compute the fast Fourier transform.

In addition to choosing the NR waveforms for the training set,
one needs to choose a relevant PSD for the mismatch.
We have opted to use a flat PSD for the mismatch calculation,
as it provides results that are independent of the detector sensitivity and mass scale.
The use of a flat PSD ensures that the improvement in accuracy is due mainly
to the difference in high-dimensional fitting.
Additionally, we are interested in examining the effect of introducing a detector PSD
on the optimization process. For this, we have chosen the
zero-detuned high-power (\zdethp) noise PSD \citep{aasi2015advanced}.
Since the total mass of the system scales with the frequency of the waveform,
we must choose a corresponding mass scale to match the frequency range of our noise PSD.
To demonstrate the effect of introducing a detector specific PSD,
we selected an arbitrary mass scale of $M=50M_{\odot}$, as binaries
of this mass are commonly observed by the LIGO and Virgo detectors~\citep{gwtc1, gwtc2, gwtc21, gwtc3}.

@@

With the loss function evaluated, we apply gradient descent to optimize the tunable
coefficients as shown in Algorithm~\ref{alg:gradient}. We take $\lambda_i$ to be the 
original coefficients given in \citep{khan2016frequency} because they likely lie in the
neighborhood of the minimum that we wish to find. 
We fix our learning rate, $\alpha$, to be $10^{-6}$, which is small enough to ensure we 
don't move far from the minimum.
Finally, we stop the optimization when the validation loss stops decreasing 
(see Sec.~\ref{sec:result} for description of the validation set).
This can be seen in Fig.~\ref{fig:loss} at around $12000$ iterations.

\begin{figure}
	\script{loss.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/loss.pdf}
	\caption{Loss functions against the number of iterations.
	The vertical dotted line indicates minimum of the validation loss,
	where we stop the optimization and obtain a new set of waveform coefficients, $\lambda$.}
	\label{fig:loss}
\end{figure}

@@

\begin{figure}[t]
	\script{intrin_space.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/intrin_space.pdf}
	\caption{Distribution of training and validation NR waveforms in the $q-\chi_{PN}$
	space. The blue stars represent the waveforms used to compute the training loss and the orange
	dots represent the waveforms used for validation.\kw{Change the legend, and include the extra points. Change the training point to be star and testing point to be dot.}}
	\label{fig:intrin_space}
\end{figure}

\begin{table}[t]
	\centering
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
		\midrule\midrule SXS:BBH:0156 & 1.0 & -0.95    & -0.95    \\
		SXS:BBH:0151 & 1.0 & -0.60    & -0.60    \\
		SXS:BBH:0001 & 1.0 &  0.00    &  0.00    \\
		SXS:BBH:0152 & 1.0 &  0.60    &  0.60    \\
		SXS:BBH:0172 & 1.0 &  0.98    &  0.98    \\
		SXS:BBH:1418 & 4.0 & -0.40    & -0.50    \\
		SXS:BBH:0167 & 4.0 &  0.00    &  0.00    \\
		SXS:BBH:1417 & 4.0 &  0.40    &  0.50    \\
		SXS:BBH:0064 & 8.0 & -0.50    & -0.46    \\
		SXS:BBH:0063 & 8.0 &  0.00    &  0.00    \\
		SXS:BBH:0065 & 8.0 &  0.50    &  0.46    \\ \midrule\bottomrule
	\end{tabularx}
	\caption{List of waveforms used to recalibrate the model. The mass ratio is defined as
	$q=m_1/m_2\geq 1$ and the spins are denoted by $\chi_{1,2}$. Out of the 11 waveforms listed
	here, nine of them are also used in the original IMRPhenomD calibration. The remaining two
	are SXS analogues to two of the BAM NR simulations used in the original calibration
	\citep{khan2016frequency}.}
	\label{tab:q148}
\end{table}
\begin{table}[t]
	\centering
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
		\midrule\midrule SXS:BBH:0234 & 2.0 & -0.85    & -0.85    \\
		SXS:BBH:0235 & 2.0 & -0.60    & -0.60    \\
		SXS:BBH:0169 & 2.0 & 0.00     & 0.00     \\
		SXS:BBH:0256 & 2.0 & 0.60     & 0.60     \\
		SXS:BBH:0257 & 2.0 & 0.85     & 0.85     \\ \midrule\bottomrule
	\end{tabularx}
	\caption{Additional NR waveforms used in further recalibration.}
	\label{tab:q1248}
\end{table}

@@

To evaluate how well the optimization procedure generalizes to waveforms that are
not in the training set, we evaluate the mismatch between the fine-tuned model
and an additional 536 NR waveforms in the SXS catalog i.e., the validation set.
We select waveforms that share the same part of the parameter space with the training
set i.e., waveforms with negligible eccentricity (${e<2\times10^{-3}}$)
and precession (${\chi_{x,y}<5\times10^{-3}}$). Figure~\ref{fig:intrin_space}
shows how the training and validation waveforms are distributed in the
$q-\chi_{\rm PN}$ space.

To illustrate the effect of optimiziation on an individual waveform level, in Fig.
\ref{fig:0154} we plot the phase and amplitude of a particular waveform before and after optimization together with
the NR waveform taken directly from the SXS catalog. In the bottom panel one can see that,
compared to the original IMRPhenomD waveform, the optimized waveform has smaller residuals both in amplitude and
phase, particularly in the inspiral region where the amplitude displays a
$50\%$ reduction in error. For a fair comparison, we selected the SXS:BBH:0154 NR waveform
which was also used in \citep{khan2016frequency} to validate the original waveform.

With the purpose of improving downstream tasks such as parameter estimation in
mind, the more relevant metric of improvement is the distribution of improvement
in mismatch over the entire validation dataset. Figure~\ref{fig:q148} shows the
distribution of log-mismatches for the validation waveforms before and after the
optimization procedure. Here we show results using a constant PSD in our loss
function. One can see the distribution of the optimized waveform is skewed toward lower mismatches,
with the peak of the distribution being shifted by approximately an order of
magnitude, and the median mismatch being reduced by 50\% (see vertical dotted lines). When using
$\mathcal{L}_{\mathrm{nl}}$, we observe a less pronounced improvement with a 22.9\%
decrease in the median.

@@

Note that the performace of the IMRPhenomD model was initally tested using the
{\zdethp} weighted mismatch.\footnote{
	Note, however, that the mismatch was never directly used during the calibration process~\citep{khan2016frequency, khan2019phenomenological}.
} We would therefore like to examine whether using the {\zdethp} PSD in our loss function
could lead to an improved to mismatch. Peforming the optimization again, we find no significant difference
between the results using the two PSDs in the distribution of mismatches.

@@

To investigate the performance of recalibration over the source parameter space,
we plot the improvement of the log-mismatch as a function of the parameter space
$q-\chi_{\mathrm{PN}}$ in Fig. \ref{fig:ps_q148_qchi}. Red points indicate that the waveform
is improved by the optimization procedure, while blue points indicate that the waveform mismatch
\textit{increases} during optimization.
We can see that waveforms with $q\le4$ show the most consistent average improvement.
This is likely due to the better coverage of training waveforms in that part of the parameter space (see Fig.~\ref{fig:intrin_space}).
On the spin axis, we can see that the waveforms with
$\chi_{\textrm{PN}}\sim0$ show the most consistent improvement. 
When we move away from $\chi_{\textrm{PN}}\sim0$, the improvement fluctuates but exhibits an overall trend. 
This is particularly true in the $q\le4$ region, where we see a consistent
improvement of the waveform for $\chi_{\textrm{PN}}<0$. 
Since the original model was developed for aligned-spin systems, NR waveforms
with opposite sign $\chi_{1,2}$ are likely less well-fit \kw{Verify this.}.
To show this, we plot the parameter space $\chi_1 - \chi_2$ in Fig.~\ref{fig:ps_q148}. 
Points along the diagonal axis, $\chi_1\sim\chi_2$, show good mismatch improvements as discussed
above. Meanwhile, the top-left ($\chi_1<0<\chi_2$) and bottom-right
($\chi_1>0>\chi_2$) regions respond to the optimization differently. In the top-left
region, the waveform generally improves after optimization. However,
in the bottom-right region, the waveform does not improve after optimization.
